---
title: "hw5"
author: "Allegra Frickel"
date: "4/4/2019"
output: html_document
---

```{r}
if (file.exists("alcohol.rds")) {
  data <- readRDS("alcohol.rds")
} else {
  url <- "https://xdaiisu.github.io/ds202materials/data/ames.csv"
  data <- readr::read_csv(url)
  saveRDS(data, file="alcohol.rds")
} 
```

**2. Data Cleaning**

Extract geographic latitude and longitude from the variable Store Location.

```{r}
library(dplyr)
library(tidyverse)
head(data$`Store Location`)
dat <- data %>% separate(`Store Location`, into = c("street", "city and zip", "coordinates"), sep = "\n") %>% separate(coordinates, into = c("lat", "long"), sep = ",") #%>% separate(prelat, into = c("(" "lat"), sep = "(")
```

Check variable types. Pick five variables that need to be converted to a different type and fix those.

```{r}
class(data$`Invoice/Item Number`)
class(data$Date)
class(data$`Store Number`)
class(data$`Store Name`)
class(data$Address)
class(data$City)
class(data$`Zip Code`) 
as.character(data$`Zip Code`)
class(data$`Store Location`)
class(data$`County Number`)
class(data$County)
class(data$Category)
as.numeric(data$Category)
class(data$`Category Name`)
class(data$`Vendor Number`)
as.numeric(data$`Vendor Number`)
class(data$`Vendor Name`)
class(data$`Item Number`)
as.character(data$`Item Number`)
class(data$`Item Description`)
class(data$Pack)
as.character(data$Pack)
class(data$`Bottle Volume (ml)`)
class(data$`State Bottle Cost`)
class(data$`State Bottle Retail`)
class(data$`Bottles Sold`)
class(data$`Sale (Dollars)`)
class(data$`Volume Sold (Liters)`)
class(data$`Volume Sold (Gallons)`)
```

Extract year, month and day from the variable Date.

```{r}
library(lubridate)
#mdy(data$Date) or...
dat <- dat %>% separate(Date, into = c("month", "day", "year"), sep = "/")
dat
```


**3. First Overview**

Plot a scatterplot of lat and long of store locations.
```{r}
library(ggplot2)
ggplot(dat, aes(x=lat, y=long)) + geom_point()
```

Provide a visual breakdown of the liquor category (by Category Name). Include volume sold in the breakdown.
```{r}
library(plotly)
plt <- ggplot(dat, aes (x=`Category Name`, color=`Volume Sold (Liters)`)) + geom_bar()
ggplotly(plt)
```

**4. Find daily liquor sales in Ames in 2018.**

Summarize number of sales, volume of liquor sold and amount of money spent. Plot volume sold by day (use a scatterplot of volume by day and facet by month). Describe any patterns. Find the dates for ISU football home games in Fall 2018. Can you see a pattern?

From the graphs below, I see some rather sparatic patterns in the plots. Perhaps this is to represent purchasing high volumes of alcohol on or for the weekends and around the holiday season. It is difficult to conclude any patterns during Fall 2018 as there is no data for it.

```{r}
df <- dat %>% filter(year==2018) %>% select(day, month, `Volume Sold (Liters)`, `Sale (Dollars)`)
summary(df)
plt <- ggplot(df, aes(x=day, y=`Volume Sold (Liters)`)) + geom_point() + facet_wrap(~month)
ggplotly()
```


